Figure supplements

### 2023-01-10 text/Manuscript.Rmd
edit figure legends

![**Figure 3. Larvae with inhibited NOS have a delayed response to UV-avoidance.** **(A)** Top: The domain and Exon/Intron structure of NOS. Bottom: Close-up region showing the genomic locus of NOS gene and the wild-type sequence (WT) targeted by CRISPR/Cas9. The generated mutants (NOSΔ11, NOSΔ23) are also shown. Pink indicates target sites. Gray shows PAM sequences, red shows stop codons. **(B)** Overlaid trajectories for WT (left, n=20) and NOS mutant (right, n=31) 3-day-old larvae. Grey shows the trajectories before UV stimulation and purple shows the trajectories after UV stimulation. **(C)** The temporal changes in the vertical position of the WT and mutant larvae before and after UV stimulation are shown. The starting points of each larval trajectory are set to 0. After UV stimulation is indicated by purple squares. **(D)** Vertical swimming in wild-type (WT) and mutant (NOSΔ11 and NOSΔ23) larvae at 3-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top. The data are shown in 30 s bins. **(E)** Vertical swimming in larvae treated with NOS inhibitors, L-NAME at 3-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top. The data are shown in 30 s bins. .](../figures/Fig3.png){width="100%"}

@@

![**Figure 4. cPRC-specific expression of NIT-GC1 localises in close proximity to NOS in the ANS region.** (D and E) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in WT (dark yellow) and NOS mutant (D: NOSΔ11/Δ11, E: NOSΔ23/Δ23) (light yellow) larvae. (A) Expression analysis of the NIT-GC1 gene (magenta) using in situ HCR. White arrowheads indicate cells expressing NIT-GC1.Co-staining image using acetylated α-tubulin antibody (ac α-tub: green). (B) Co-expression analysis image of the c-opsin1 gene (green). Black arrowheads indicate cells expressing c-opsin1 differently from cPRC. (C and D) Localisation analysis using NIT-GC1 and NIT-GC2 antibodies. Green shows co-staining with acetylated α-tubulin antibody (ac α-tub).(E-G) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in control (non-injection) and NIT-GC1 and NIT-GC2 knockdown by morpholino (MO1 and 2) larvae.(H) The domain structure of NIT-GC1. Grey squares indicate the transmembrane domain.(I) A schematic of the NIT-GC1 assay; COS7 cells were used to express NIT-GC1 and the cGMP indicator Green cGull, and the NO donor SNAP and DMSO as control. Green cGull shows cGMP-dependent green fluorescence. (J-M) The intensity changes in cGMP fluorescence over time for the four conditions are shown. Gray traces show the optical responses to individual intensities (n \> 6). The individual intensities are normalized (ΔF/F0). The dark line shows the mean value and the band the standard error range of the mean. The respective solution was acted on 2 min after the start of imaging (grey bars).](../figures/Fig4.png){width="100%"}

We next quantified the responses of the INNOS and INRGW interneurons following a UV/violet stimulus to the cPRCs. To unambiguously identify the cells from which we recorded calcium signals, we developed an on-slide immunostaining method. We used the cell-specific neuropeptide markers anti-RYamide (INNOS) and anti-RGWa (INRGW) [@Conzelmann2011] to mark the interneurons. Based on the position of the nuclei, we could correlate live and fixed samples at a single-cell precision (Figure 4---figure supplement 2). In both wild type and *NOS*-mutant larvae, INNOS cells showed increase calcium signals during the inhibition phase of the cPRCs (Figure 4D and E). In addition, In-NOS depolarisation was suppressed in NIT-GC2 molphants where cPRC hyperpolarisation was suppressed. This indicates that cPRC hyperpolarisation is required for IN-NOS depolarisation (Figure 4F). In contrast, calcium signals increased in INRGW cells in wild-type larvae during the second activation phase of cPRCs, a response that was lacking in *NOS*-mutants (Figure 4G and H). This suggests that there is no circuit output to the Ser-h1 cells and the ciliary band in *NOS*-mutants, consistent with the observed behavioural defects(Figure 4I).

>>>>

![**Figure 3. Larvae with inhibited NOS have a delayed response to UV-avoidance.** **(A)** Top: The domain and Exon/Intron structure of NOS. Bottom: Close-up region showing the genomic locus of NOS gene and the wild-type sequence (WT) targeted by CRISPR/Cas9. The generated mutants (NOSΔ11, NOSΔ23) are also shown. Pink indicates target sites. Gray shows PAM sequences, red shows stop codons. **(B)** Overlaid trajectories for WT (left, n=20) and NOS mutant (right, n=31) 3-day-old larvae. Grey shows the trajectories before UV stimulation and purple shows the trajectories after UV stimulation. **(C)** The temporal changes in the vertical position of the WT and mutant larvae before and after UV stimulation are shown. The starting points of each larval trajectory are set to 0. After UV stimulation is indicated by purple squares. **(D)** Vertical swimming in wild-type (WT) and mutant (NOSΔ11 and NOSΔ23) larvae at 3-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top. The data are shown in 30 s bins. **(E)** Vertical swimming in larvae treated with NOS inhibitors, L-NAME at 3-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top. The data are shown in 30 s bins.](../figures/Fig3.png){width="100%"}

@@

![**Figure 4. cPRC-specific expression of NIT-GC1 localises in close proximity to NOS in the ANS region.** **(A, B)** Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in WT (dark yellow) and NOS mutant (A: NOSΔ11/Δ11, B: NOSΔ23/Δ23) (light yellow) larvae. **(C, D)** Expression analysis of the NIT-GC1 (C) and NIT-GC2 (D) gene (magenta) using in situ HCR. Co-staining image using acetylated α-tubulin antibody (ac α-tub: green). **(E, F)** Localisation analysis using NIT-GC1 and NIT-GC2 antibodies. Green shows co-staining with acetylated α-tubulin antibody (ac α-tub). **(G, H)** Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in NIT-GC1 (G) and NIT-GC2 (H) knockdown by morpholino (MO1 and 2) larvae. **(I)** The domain structure of NIT-GC1. Grey squares indicate the transmembrane domain. **(J)** A schematic of the NIT-GC1 assay; COS7 cells were used to express NIT-GC1 and the cGMP indicator Green cGull, and the NO donor SNAP and DMSO as control. Green cGull shows cGMP-dependent green fluorescence. **(K-N)** The intensity changes in cGMP fluorescence over time for the four conditions are shown. Gray traces show the optical responses to individual intensities (n \> 6). The individual intensities are normalized (ΔF/F0). The dark line shows the mean value and the band the standard error range of the mean. The respective solution was acted on 2 min after the start of imaging (grey bars).](../figures/Fig4.png){width="100%"}

We next quantified the responses of the INNOS and INRGW interneurons following a UV/violet stimulus to the cPRCs. To unambiguously identify the cells from which we recorded calcium signals, we developed an on-slide immunostaining method. We used the cell-specific neuropeptide markers anti-RYamide (INNOS) and anti-RGWa (INRGW) [@Conzelmann2011] to mark the interneurons. Based on the position of the nuclei, we could correlate live and fixed samples at a single-cell precision (Figure 4---figure supplement 2). In both wild type and *NOS*-mutant larvae, INNOS cells showed increase calcium signals during the inhibition phase of the cPRCs (Figure 4D and E). In addition, In-NOS depolarisation was suppressed in NIT-GC2 molphants where cPRC hyperpolarisation was suppressed. This indicates that cPRC hyperpolarisation is required for IN-NOS depolarisation (Figure 4F). In contrast, calcium signals increased in INRGW cells in wild-type larvae during the second activation phase of cPRCs, a response that was lacking in *NOS*-mutants (Figure 4G and H). This suggests that there is no circuit output to the Ser-h1 cells and the ciliary band in *NOS*-mutants, consistent with the observed behavioural defects (Figure 4I).

### 2023-01-10 text/Supplements.Rmd
edit figure legends

@@

![**Figure 4---figure supplement 2.** (A) Immunostaining with NOS antibodies. (B) Co-expression analysis image with the RYamide proneuropeptide gene (RYa-pNP: green), which is highly expressed in INNOS. White arrowheads indicate co-expressed INNOS. Black arrowheads indicate larval eyes. White dashed lines indicate the outer frame of the larva. (A) Co-expression analysis image of the MLD-pedal2 amide proneuropeptide gene (MLD-pNP: green). Black arrowheads indicate cells different from cPRC. (B) Co-localisation analysis using NIT-GC1 and NOS antibodies.](../figures/Fig4_sup2.png){width="100%"}

>>>>

![**Figure 1---figure supplement 2.** Expression analysis of the NOS gene (magenta) using in situ HCR using the larva at 3-day-old in dorsal view (A) and anterior view (B). (B) Insert: showing two NOS cells on each side.](../figures/Fig1_sup2.png){width="100%"}

![**Figure 3---figure supplement 1.** **(A)** Overlaid trajectories for WT (left, n=20) and NOS mutant (right, n=31) 2-day-old larvae. Grey shows the trajectories before UV stimulation and purple shows the trajectories after UV stimulation. **(B)** The temporal changes in the vertical position of the WT and mutant 2-days-old larvae before and after UV stimulation are shown. The starting points of each larval trajectory are set to 0. After UV stimulation is indicated by purple squares. **(C)** Vertical swimming in wild-type (WT) and mutant (NOSΔ11 and NOSΔ23) larvae at 2-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top. The data are shown in 30 s bins. **(D, E)** The temporal changes in the distance traveled of the WT and mutant in 2-day (D) and 3-day-old (E) larvae before and after UV stimulation are shown.](../figures/Fig3_sup1.png){width="100%"}

@@

![**Figure 4---figure supplement 2.** **(A)** Co-expression analysis image of the NIT-GC1 (magenta) and MLD-pedal2 amide proneuropeptide gene (MLD: green). Anterior view of the larva at 2-day-old. **(B, C)** Expression analysis of the NIT-GC2 gene (magenta) using in situ HCR. Anterior (B) and posterior (C) views of the larva at 3-day-old. **(D, E)** Localisation analysis for NIT-GC1 (B) and NIT-GC2 (C) morphant using NIT-GC1 and NIT-GC2 antibodies. Green shows co-staining with acetylated α-tubulin antibody (acTub). Anterior view of the larva at 2-day-old. **(F)** Co-localisation analysis using NIT-GC1 (magenta) and NOS (green) antibodies. Anterior view of the larva at 2-day-old.](../figures/Fig4_sup2.png){width="100%"}
